Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-02-17 main.tex
Fix ESOP'18 broken link
From Lars' homepage, as it is no longer available in Lau's.
Checkpoint: adjust existing terminology

\leo{We might need to change the intuition behind some of these things.}
A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow
\{\HIGHSEC,\LOWSEC\} \times \{\HIGHINT, \LOWINT\}$\bcp{We need more mnemonic names for
  these (that help us keep straight when we're talking about confidentiality
  and when integrity) in the way they are displayed.}
is a map from machine components
to confidentiality and integrity levels. $\LOWSEC$ marks locations whose
contents are observable from the outside (by the ``attacker,'' e.g.,
by DMA) and $\HIGHSEC$ marks internal ``hidden'' state that cannot be
observed directly (e.g., processor registers and memory addresses for
temporary variables will usually be hidden). Similarly, $\LOWINT$
locations correspond to low-integrity components (e.g. locations that
can be written to \iftext \bcp{directly or indirectly} \fi by the network), while $\HIGHINT$ locations should be
protected from such flows.

We denote this (reflexive) flows-to relation with $\flows$. Then
$\LOWSEC \flowsstrict \HIGHSEC$ and $\HIGHINT \flowsstrict
\LOWINT$. We can easily lift this relation to pairs of secrecy and
integrity labels, and from there to contours pointwise.

@@

@@

In particular, the observation $\obstrace(\contour, \seq{\mach})$
arising from a pair of a contour $\contour$ and a trace $\seq{\mach}$
is sequence of state observations $\stateobs_0,\, \stateobs_1,\,
\ldots$ defined informally as follows:
\begin{itemize}
\item $\stateobs_0$ maps every visible component $\component$ (where
  $\contour(\component) = (\LOWSEC, \_)$) to its value in $\mach_0$
  and every hidden component to $\bot$;
\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for
  each visible component $\component$ (where $\contour(\component) =
  (\LOWSEC, \_)$), copying its value from $\mach_{i+1}$, and (2) for
  each hidden component $\component$, where $\contour (\component) =
  (\HIGHSEC,\_)$, copying its value from $\stateobs_{i}$, and (3) if
  the resulting $\stateobs_{i+1}$ is identical to $\stateobs_{i}$,
  then skipping $\stateobs_{i+1}$.  \iftext\bcp{Too informal!!}\fi
\end{itemize}

@@

Given a machine state $\mach$ and a contour $\contour$, a {\em
  variant} of $\mach$ with respect to $\contour$ is any machine state
$\mach'$ whose components agree with $\mach$ on all components marked
$\LOWSEC$ by $\contour$---i.e., if $\contour(\component) = \LOWSEC$, then
$\mach(\component) = \mach'(\component)$.

@@

A trace $\seq{M}$ satisfies the {\em integrity} safety property with
respect to a contour $\contour$, if for all components $\component$
where $\contour(\component) = (\_, \HIGHINT)$, $M_i(K) = M_0(K)$ for
all $i$ in the trace.

@@

\item StkTokens: Skorstengaard, Devriese, Birkedal
\begin{itemize}
\item POPL \cite{skorstengaard2019stktokens}
\item tech report with proofs: \cite{skorstengaard2018stktokens}
\url{https://arxiv.org/pdf/1811.02787.pdf}
\item TR with proofs for ``Reasoning about a machine...'', Feb 2019:
\url{https://arxiv.org/pdf/1902.05283.pdf}
\item \url{http://cs.au.dk/~lask/papers/esop2018.pdf}
\item Also: {\em Reasoning about a Machine with Local Capabilities: Provably
  Safe Stack and Return Pointer Management} \cite{10.1145/3363519}
\url{https://dl.acm.org/doi/fullHtml/10.1145/3363519}

>>>>

A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow \{\ACCYES,
\ACCNO, \ACCCODE\}$ is a map from machine components to accessibility levels.
$\ACCYES$ marks locations whose contents represent data observable from the
outside (by the ``attacker,'' e.g., by DMA), which moreover correspond to
low-integrity components (e.g., locations that can be written to \iftext
\bcp{directly or indirectly} \fi by the network). $\ACCNO$ marks internal
``hidden'' state that cannot be observed directly (e.g., processor registers and
memory addresses for temporary variables will usually be hidden), which is
moreover high-integrity (i.e., locations which should be protected from flows
from the network)
%
\rb{Tried to merge discussion on integrity and secrecy, but maybe it is clearer
  to separate both concerns?}
%
In addition to those two accessibility levels, $\ACCCODE$ marks locations that
hold executable instructions, which are considered to be low-secrecy (i.e.,
observable by the attacker) and high-integrity (i.e., the attacker cannot write
to them).

\rb{OLD:
\leo{We might need to change the intuition behind some of these things.}
A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow
\{\HIGHSEC,\LOWSEC\} \times \{\HIGHINT, \LOWINT\}$\bcp{We need more mnemonic names for
  these (that help us keep straight when we're talking about confidentiality
  and when integrity) in the way they are displayed.}
is a map from machine components
to confidentiality and integrity levels. $\LOWSEC$ marks locations whose
contents are observable from the outside (by the ``attacker,'' e.g.,
by DMA) and $\HIGHSEC$ marks internal ``hidden'' state that cannot be
observed directly (e.g., processor registers and memory addresses for
temporary variables will usually be hidden). Similarly, $\LOWINT$
locations correspond to low-integrity components (e.g. locations that
can be written to \iftext \bcp{directly or indirectly} \fi by the network), while $\HIGHINT$ locations should be
protected from such flows.
}

\rb{The flows-to relation was quite nice before, now it looks like it will get
  messier\ldots}
%
We denote this (reflexive) flows-to relation with $\flows$. Then
$\LOWSEC \flowsstrict \HIGHSEC$ and $\HIGHINT \flowsstrict
\LOWINT$. We can easily lift this relation to pairs of secrecy and
integrity labels, and from there to contours pointwise.

@@

\rb{We are not using observations at the moment, but a few adjustments are made
  here.}

@@

In particular, the observation $\obstrace(\contour, \seq{\mach})$
arising from a pair of a contour $\contour$ and a trace $\seq{\mach}$
is sequence of state observations $\stateobs_0,\, \stateobs_1,\,
\ldots$ defined informally as follows:
\begin{itemize}
\item $\stateobs_0$ maps every visible component $\component$ (where
  $\contour(\component) \in \{\ACCYES, \ACCCODE\}$) to its value in $\mach_0$
  and every hidden component to $\bot$;
\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for
  each visible component $\component$ (where $\contour(\component) \in
  \{\ACCYES, \ACCCODE\}$), copying its value from $\mach_{i+1}$, and (2) for
  each hidden component $\component$, where $\contour (\component) =
  \ACCNO$, copying its value from $\stateobs_{i}$, and (3) if
  the resulting $\stateobs_{i+1}$ is identical to $\stateobs_{i}$,
  then skipping $\stateobs_{i+1}$.  \iftext\bcp{Too informal!!}\fi
\end{itemize}

@@

Given a machine state $\mach$ and a contour $\contour$, a {\em
  variant} of $\mach$ with respect to $\contour$ is any machine state
$\mach'$ whose components agree with $\mach$ on all components marked
$\ACCYES$ or $\ACCCODE$ by $\contour$---i.e., if $\contour(\component) \in \{\ACCYES, \ACCCODE\}$, then
$\mach(\component) = \mach'(\component)$.
%
\rb{When we talk about observations, it is visibility that we are concerned with
  and moving from $\HIGHSEC$ to $\ACCNO$ and $\LOWSEC$ to $\ACCYES$ and
  $\ACCCODE$ is simple. Here, the case of code is slightly more
  interesting (are $\ACCCODE$-variants ever useful?).}

@@

A trace $\seq{M}$ satisfies the {\em integrity} safety property with
respect to a contour $\contour$, if for all components $\component$
where $\contour(\component) = (\_, \HIGHINT)$, $M_i(K) = M_0(K)$ for
all $i$ in the trace.
%
\rb{Here, $\HIGHINT$ would be replaced by $\in \{\ACCNO, \ACCCODE\}$.}

@@

\item StkTokens: Skorstengaard, Devriese, Birkedal
\begin{itemize}
\item POPL \cite{skorstengaard2019stktokens}
\item tech report with proofs: \cite{skorstengaard2018stktokens}
\url{https://arxiv.org/pdf/1811.02787.pdf}
\item TR with proofs for ``Reasoning about a machine...'', Feb 2019:
\url{https://arxiv.org/pdf/1902.05283.pdf}
\item \url{https://cs.au.dk/~birke/papers/local-capabilities-conf.pdf}
\item Also: {\em Reasoning about a Machine with Local Capabilities: Provably
  Safe Stack and Return Pointer Management} \cite{10.1145/3363519}
\url{https://dl.acm.org/doi/fullHtml/10.1145/3363519}
